---
title: "Assignments"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

This page will contain all the assignments you submit for the class.



### Instructions for all assignments

I want you to submit your assignment as a PDF, so I can keep a record of what the code looked like that day. I also want you to include your answers on your personal GitHub website. This will be good practice for editing your website and it will help you produce something you can keep after the class is over.

1. Download the Assignment1.Rmd file from Canvas. You can use this as a template for writing your answers. It's the same as what you can see on my website in the Assignments tab. Once we're done with this I'll edit the text on the website to include the solutions.

2. On RStudio, open a new R script in RStudio (File > New File > R Script). This is where you can test out your R code. You'll write your R commands and draw plots here.

3. Once you have finalized your code, copy and paste your results into this template (Assignment 1.Rmd). For example, if you produced a plot as the solution to one of the problems, you can copy and paste the R code in R markdown by using the ` ``{r} ``` ` command. Answer the questions in full sentences and Save.

4. Produce a PDF file with your answers. To do this, knit to PDF (use Knit button at the top of RStudio), locate the PDF file in your docs folder (it's in the same folder as the Rproj), and submit that on on Canvas in Assignment 1.

5. Build Website, go to GitHub desktop, commit and push. Now your solutions should be on your website as well.



# Assignment 1

**Collaborators: Rachael Villari, Michaela Reiser. **

This assignment is due on Canvas on Monday 9/20 before class, at 10:15 am. Include the name of anyone with whom you collaborated at the top of the assignment.

Collaborators: Rachael Villari, Michaela Reiser 


### Problem 1 

Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.

```{r}
library(datasets)
```

Load the USArrests data set and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?

Answer: It is useful to rename the dataset so we have a single, shorter variable to account for the whole dataset

```{r}
dat <- USArrests
```

### Problem 2

Use this command to make the state names into a new variable called State. 

```{r, eval=FALSE}
dat$state <- tolower(rownames(USArrests))
```

This dataset has the state names as row names, so we just want to make them into a new variable. We also make them all lower case, because that will help us draw a map later - the map function requires the states to be lower case.


List the variables contained in the dataset `USArrests`.

Answer: Murder, Assault, Urban Population, Rape

```{r}
names(dat)
```

### Problem 3 

What type of variable (from the DVB chapter) is `Murder`? 

Answer: quantitative

What R Type of variable is it?

Answer: numeric


### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

Answer: The numbers are the frequency of arrests per crime in each state and the urban population is the percentage of people
in each state who live in an urban area. The information in general is the US Arrests of each crime in each state.

### Problem 5

Draw a histogram of `Murder` with proper labels and title.

```{r}
hist(dat$Murder, main = "Histogram of US Arrests of Murder", 
     xlab = "US Arrests of Murder", xlim = c(0,20), 
     ylim = c(0, 15))
```

### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

Answer: 
Min: 0.8 
Max:17.4
Mean:7.788
Median:7.250
Q1: 4.075
Q3: 11.250
The difference between mean and median is that the mean gives the average of all murders while the median relates to the value at the midpoint of the data set. R gives the 1st and 3rd quartiles so we can have a fuller picture of our dataset and we can have a better idea of where our median falls. 

```{r}
summary(dat$Murder)
```

### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r, echo = TRUE, fig.width = 5, fig.height = 8}
hist(dat$Assault, main = "Histogram of US Arrests of Assault", 
     xlab = "US Arrests of Assault", xlim = c(0,350), 
     ylim = c(0,15))
hist(dat$UrbanPop, main = "Histogram of US Urban Population",
     xlab = "US Urban Population", xlim = c(0,100), 
     ylim = c(0,15))
hist(dat$Rape, main = "Histogram of US Arrests of Rape", 
     xlab = "US Arrests of Rape", xlim = c(0,50), ylim = c(0,15))
par(mfrow=c(3,1))
hist(dat$Assault, main = "Histogram of US Arrests of Assault", 
     xlab = "US Arrests of Assault", xlim = c(0,350), ylim = c(0,15))
hist(dat$UrbanPop, main = "Histogram of US Urban Population",
     xlab = "US Urban Population", xlim = c(0,100), ylim = c(0,15))
hist(dat$Rape, main = "Histogram of US Arrests of Rape", 
     xlab = "US Arrests of Rape", xlim = c(0,50), ylim = c(0,15))
```

What does the command par do, in your own words (you can look this up by asking R `?par`)?

Answer: command par allows one to see all three graphs together in order to compare them more easily

What can you learn from plotting the histograms together?

Answer: You can learn the distributions against each other and see which variable has the most uniform or skewed distribution.
  
### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.

Run this code:

```{r, eval = FALSE, fig.width = 7.5, fig.height = 4}
install.packages("maps")
install.packages("ggplot2")
library('maps') 
library('ggplot2') 

ggplot(dat, aes(map_id=state, fill=Murder)) + 
  geom_map(map=map_data("state")) + 
  expand_limits(x=map_data("state")$long, y=map_data("state")$lat)
```

What does this code do? Explain what each line is doing.

Answer: The first two lines of code are installing the packages onto R so we can use the functions located within the package. Calling the library of maps and ggplot2 allows us to access these two functions within our own R Script. Lines 167-169 are plotting our USArrests data located in the variable dat, specifically the variable of "Murder." It maps the data by U.S state of the different concentrations of murder arrests.

$$\\[2in]$$




# Assignment 2 

(Coming soon)


